---
title: "Final_Project"
author: "Matteo Larrode"
date: "2023-04-10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

A growing body of literature is interested in the effect of the rise and diffusion of new technologies on international relations. One particular topic we are interested in is the relationship between internet use, especially activity on social networks, and freedom of speech.

Social media platforms have fundamentally transformed the landscape of social and political movements. They have allowed individuals to bypass traditional gatekeepers of information, such as the mainstream media, and connect with each other directly. This has enabled dissidents to spread information and organize more quickly and effectively than ever before. But on the other hand, they can also be easily monitored and censored by governments, potentially putting activists at risk.

## Data Preparation

We created a dataset called “final_data.csv” by joining data on Internet penetration, provided by the [International Telecommunication Union (ITU)](https://www.itu.int/en/ITU-D/Statistics/Pages/stat/default.aspx) (percentage of the population who has access to the internet), and around 50 different metrics used by [V-Dem](https://www.v-dem.net/data/the-v-dem-dataset/) to measure levels of democracy. This data is available for a total of 157 countries.

```{r loading, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(scales)
library(stargazer)


df <- read_csv("data/final_data.csv")

df_noNA <- read_csv("data/final_data.csv")%>% 
  drop_na()
```

```{r wrangling, message=FALSE, include=FALSE}
#dataframe for freedom of discussion, Internet penetration & Level of democracy

df_reduced <- df_noNA %>% 
  select(Economy, value_2021_perc, v2x_libdem, v2cldiscm, v2cldiscw)%>%
  mutate(is_dem = v2x_libdem >= 0.5,
         is_dem_long = ifelse(is_dem, "Higher Liberal Dem. Index", "Lower Liberal Dem. Index"))

```

## Findings

### Effect of Internet on Freedom of Speech for Women

My main research question is: “Does higher Internet penetration lead to higher levels of freedom of speech?” More specifically, could the Internet empower people who are usually less able to express themselves to finally share their opinions?

Here is a scatterplot that illustrates the relationship between freedom of discussion for women, my outcome variable, and access to the Internet, my main independent variable.

```{r freedom_speech_women, dpi= 200}
free_disc_women_plot <- ggplot(df_reduced, aes(x = value_2021_perc, y = v2cldiscw))+
  geom_point()+
  geom_smooth(method = "lm")+
  
  theme_wsj(title_family = "Roboto")+
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(hjust=0.5, face="bold", size=12, margin = margin(b=10)),
        plot.caption = element_text(size=9, vjust = -2),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -2),
        axis.text.x = element_text(vjust = -.5),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank())+
  scale_x_continuous(name ="Internet Penetration",labels = scales::percent_format(accuracy = 1))+
  scale_y_continuous(name = "Freedom of Discussion for Women")+
  labs(title = "Higher Access to the Internet is Correlated with More Freedom of Discussion for Women",
       caption="Source: ITU, V-Dem | @matteoStats")

free_disc_women_plot
```

From this simple scatterplot, and the linear regression fitted to the data, it seems that countries with higher levels of Internet penetration are associated with higher levels of freedom of discussion for women. However, an important omitted variable bias could emerge if we do not take governance into account. Indeed, democracies, usually characterized by high levels of freedom of discussion, are also often more developed than authoritarian regimes, and therefore have higher Internet penetration rates on average.

An interesting extra step to analyze the relationship between Internet penetration and freedom of discussion could be to control for the level of democracy. To reach this objective, we will use V-Dem’s Liberal democracy index (v2x_libdem), that measures to what extent is the ideal of liberal democracy achieved. To clarify, the liberal principle of democracy emphasizes the importance of protecting individual and minority rights against the tyranny of the state and the tyranny of the majority. This is achieved by constitutionally protected civil liberties, strong rule of law, an independent judiciary, and effective checks and balances that, together, limit the exercise of executive power.

To explore this addition of an independent variable to the model, we can create the same scatterplot as before, but faceted by different levels of the liberal democracy index. There is no specific threshold level at which a country is considered a democracy on the V-Dem index. Rather, the index measures the extent to which a country meets the various
criteria for democracy and assigns a score accordingly. The higher the score, the more democratic the country is considered to be. However, for the purpose of this graph, I arbitrarily decided of a threshold (0.5) to separate countries into two categories of liberal democracy.

```{r freedom_speech_women_faceted, dpi= 200}
free_disc_women_plot2 <- ggplot(df_reduced, aes(x = value_2021_perc, y = v2cldiscw))+
  geom_point()+
  geom_smooth(method = "lm")+
  
  facet_wrap(~ is_dem_long, ncol=2) +
  
  theme_wsj(title_family = "Roboto")+
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(hjust=0.5, face="bold", size=12, margin = margin(b=10)),
        plot.caption = element_text(size=9, vjust = -2),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -2),
        axis.text.x = element_text(vjust = -.5),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank())+
  scale_x_continuous(name ="Internet Penetration",labels = scales::percent_format(accuracy = 1))+
  scale_y_continuous(name = "Freedom of Discussion for Women")+
  labs(title = "Relationship between Internet Penetration and Freedom of Discussion for Women \n at Different Levels of Liberal Democracy",
       caption="Source: ITU, V-Dem | @matteoStats")

free_disc_women_plot2
```

This faceted graph gives some very interesting information! It suggests that the effect of Internet penetration on freedom of discussion for women is **conditional on the level of libertarian democracy**. In countries considered to be more democratic, higher access to the Internet seems to be correlated to an increase in the freedom of discussion for women. This observation supports the hypothesis that the inter-connectivity and anonymity provided by the Internet empowers groups that tend to be discriminated against to express themselves. On the other hand, countries classified as having a lower liberal democracy index do not display any apparent relationship between Internet penetration and freedom of discussion for women. This could be explained by the censorship and online harassment efforts put in place by authoritarian regimes to limit the danger that free speech on the Internet poses to their stability.




### Go deeper into why authocracies are flat
##### censorship + surveillance
##### violence
--> ACTORS OF CYBER VIOLENCE


### Look into difference between men & women + minorities (?)
--> Whose voices are given a platform through the internet


###Reminders
#### Quality of data
--> violence on women not reported ...




