---
title: "Final_Project"
author: "Matteo Larrode"
date: "2023-04-10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

A growing body of literature has been interested in the effect of the rise and diffusion of new technologies on international relations. One particular topic of contention is the relationship between internet use, especially activity on social networks, and civil liberties. 

Social media platforms have fundamentally transformed the landscape of social and political movements. They have allowed individuals to bypass traditional gatekeepers of information, such as the mainstream media, and connect with each other directly. This has enabled dissidents to spread information and organize more quickly and effectively than ever before. The other side of the coin is that they also enabled certain governments to better monitor and gather information, potentially putting activists at risk.

Through a combination of data analysis and case studies, this report seeks to further uncover the nature of the effect of the development of the internet and social media on freedom of speech and activism.  

## Dataset & Methodology

For this research project, a dataset on internet penetration, V-Dem indicators, and the political environment of the internet and social media was compiled for 179 countries.

- **Internet penetration**: The [International Telecommunication Union (ITU)](https://www.itu.int/en/Pages/default.aspx) is the UN specialized agency for information and communications technologies (ICT), and the official source for global ICT [statistics](https://www.itu.int/en/ITU-D/Statistics/Pages/stat/default.aspx). These statistics include the percentage of the population having access to the internet, broken down by gender and urban/rural area, at country level. 

- **Measure of democracy**: The [Varieties of Democracy Dataset, version 13](https://www.v-dem.net/data/the-v-dem-dataset/) measures democracy and many of its indicators in the set of countries studied. It also includes the Regimes of the World (RoW) variable that was used to classify regime types for this project. 

- **Political environment of the internet and social media**: Through expert-coded surveys, the Digital Society Survey, designed by the [Digital Society Project](http://digitalsocietyproject.org), provides information on topics related to coordinated information operations, digital media freedom, online media polarization, social cleavages as well as state internet regulation capacity and approach. 

```{r loading, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(scales)
library(stargazer)


df <- read_csv("data/final_data.csv")
```

#### Types of political systems

This report tackles the differences in the impact of the internet on civil liberties, especially freedom of expression, across different types of regimes. To classify them, we will use the Regimes of the World (RoW) data by political scientists Anna Lührmann, Marcus Tannenberg, and Staffan Lindberg, published by the Varieties of Democracy (V-Dem) project. The RoW data distinguishes four types of political systems based on how their elections work and other factors related to how their government functions. 

- closed autocracies
- electoral autocracies
- electoral democracies
- liberal democracies

One strength of using the RoW data is that it is based on evaluations by experts, who are primarily academics, members of the media, and civil society, and often nationals or residents of the country they assess. This allows for a nuanced and informed assessment of a country's political system. Additionally, V-Dem uses several experts per country, year, and topic, to make its assessments less subjective, which further increases the reliability of the data.

However, a weakness of using expert evaluations is that they are to some degree subjective, and there may be disagreement about specific characteristics or how something as complex as a political system can be reduced into a single measure. V-Dem addresses this by using several experts and specific questions on completely explained scales.

Overall, the use of the RoW data is a valid and [widely accepted approach](https://ourworldindata.org/regimes-of-the-world-data) to measuring political regimes, and we believe it will be a useful tool for our analysis.


```{r wrangling, message=FALSE, include=FALSE}
#dataframe for freedom of discussion, Internet penetration & Level of democracy

df_reduced <- df %>% 
  select(Economy, most_recent_perc, v2x_regime, v2x_libdem, v2cldiscm, v2cldiscw)%>%
  mutate(regime_type = ifelse(grepl("autocracy", v2x_regime), "Autocracies", "Democracies"))

```

## Findings

### Effect of Internet on Freedom of Speech for Women

Here is a scatterplot that illustrates the relationship between freedom of discussion for women, my outcome variable, and access to the Internet, my main independent variable.

```{r freedom_speech_women, dpi= 200}
free_disc_women_plot <- ggplot(df_reduced, aes(x = most_recent_perc, y = v2cldiscw))+
  geom_point()+
  geom_smooth(method = "lm")+
  
  theme_wsj(title_family = "Roboto")+
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(hjust=0.5, face="bold", size=12, margin = margin(b=10)),
        plot.caption = element_text(size=9, vjust = -2),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -2),
        axis.text.x = element_text(vjust = -.5),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank())+
  scale_x_continuous(name ="Internet Penetration",labels = scales::percent_format(accuracy = 1))+
  scale_y_continuous(name = "Freedom of Discussion for Women")+
  labs(title = "Higher Access to the Internet is Correlated with More Freedom of Discussion for Women",
       caption="Source: ITU, V-Dem | @matteoStats")

free_disc_women_plot
```

From this simple scatterplot, and the linear regression fitted to the data, it seems that countries with higher levels of Internet penetration are associated with higher levels of freedom of discussion for women. However, an important omitted variable bias could emerge if we do not take governance into account. Indeed, democracies, usually characterized by high levels of freedom of discussion, are also often more developed than authoritarian regimes, and therefore have higher Internet penetration rates on average.

An interesting extra step to analyze the relationship between Internet penetration and freedom of discussion could be to control for the level of democracy. To reach this objective, we will use V-Dem’s Liberal democracy index (v2x_libdem), that measures to what extent is the ideal of liberal democracy achieved. To clarify, the liberal principle of democracy emphasizes the importance of protecting individual and minority rights against the tyranny of the state and the tyranny of the majority. This is achieved by constitutionally protected civil liberties, strong rule of law, an independent judiciary, and effective checks and balances that, together, limit the exercise of executive power.

To explore this addition of an independent variable to the model, we can create the same scatterplot as before, but faceted by different levels of the liberal democracy index. There is no specific threshold level at which a country is considered a democracy on the V-Dem index. Rather, the index measures the extent to which a country meets the various
criteria for democracy and assigns a score accordingly. The higher the score, the more democratic the country is considered to be. However, for the purpose of this graph, I arbitrarily decided of a threshold (0.5) to separate countries into two categories of liberal democracy.

```{r freedom_speech_women_faceted, dpi= 200}
free_disc_women_plot2 <- ggplot(df_reduced, aes(x = most_recent_perc, y = v2cldiscw))+
  geom_point()+
  geom_smooth(method = "lm")+
  
  facet_wrap(~ regime_type, ncol=2) +
  
  theme_wsj(title_family = "Roboto")+
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(hjust=0.5, face="bold", size=12, margin = margin(b=10)),
        plot.caption = element_text(size=9, vjust = -2),
        axis.title = element_text(size = 12),
        axis.title.x = element_text(vjust = -2),
        axis.text.x = element_text(vjust = -.5),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank())+
  scale_x_continuous(name ="Internet Penetration",labels = scales::percent_format(accuracy = 1))+
  scale_y_continuous(name = "Freedom of Discussion for Women")+
  labs(title = "Relationship between Internet Penetration and Freedom of Discussion for Women \n at Different Levels of Liberal Democracy",
       caption="Source: ITU, V-Dem | @matteoStats")

free_disc_women_plot2
```

This faceted graph gives some very interesting information! It suggests that the effect of Internet penetration on freedom of discussion for women is **conditional on the level of libertarian democracy**. In countries considered to be more democratic, higher access to the Internet seems to be correlated to an increase in the freedom of discussion for women. This observation supports the hypothesis that the inter-connectivity and anonymity provided by the Internet empowers groups that tend to be discriminated against to express themselves. On the other hand, countries classified as having a lower liberal democracy index do not display any apparent relationship between Internet penetration and freedom of discussion for women. This could be explained by the censorship and online harassment efforts put in place by authoritarian regimes to limit the danger that free speech on the Internet poses to their stability.


## Two opposing effects

### Positive: the Internet & Social Media as a Promoter of Civil Liberties
#### Platform for people to speak out + gendered approach 
#### Platform to organize offline action

### Negative: Risks for democracy + gendered approach
#### i- surveillance + violence
#### ii - personalisation
#### iii - disinformation


> COMPARE AUTOCRACIES TO DEMOCRACIES = uncover nature interaction regime x internet on freedom of speech
  - Literature
  - Dataviz & models


###Reminders
#### Quality of data
--> violence on women not reported ...




